Adding selection to \stdpopsim: oh how the tables have turned

### 2024-05-23 analysis2_manuscript.tex

\section*{Introduction}
    \label{introduction}
As the complexity of population genetic datasets has increased,
simulation has become a critical tool for interpretation, 
analysis, and exploration of realistic evolutionary models.
Simulation in population genetics has a long history 
including both backward in time coalescent simulations
\citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
and forward in time simulations of complex demography and selection
\citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.
The development of simulation tools has been driven by the need to
understand the effects of complex evolutionary processes on genetic
variation, to provide a null model for hypothesis testing, to
to explore the power and limitations of statistical methods, and 
increasingly to provide a basis for machine learning and other
simulation-based inference methods.

A lingering challenge with simulation in population genetics has been
reproducibility and the ability to share and compare results among 
researchers. This challenge has been addressed in part by the development
of community resources for sharing and distributing simulation software
via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
provides a standardized interface for accessing a wide range of
population genetic models, and has begun to be be widely adopted by the community.
While that is so, the current version of \stdpopsim does not include
models of selection, which is a major limitation for many applications
of population genetic simulation. In particular, modeling selection
through simulation is critical for understanding the processes such
as adaptation, the effects of selective sweeps, and the impact of
background selection on genetic diversity. Ideally, one would like
to have a single, unified framework for simulating both neutral and
non-neutral evolutionary processes, and to be able to compare the
results of these simulations to empirical data in a manner that is
accessible to a wide range of researchers and highly reproducible. 
Further the framework should include the complex realities of 
genomes, including heterogenous recombination rates, 
variation in the density of functional elements, and relevant
population size histories.

>>>>

\section*{Introduction}
    \label{introduction}
    % natural selection
    Natural selection is a fundamental force in evolution, shaping the
    genetic diversity of populations and driving the adaptation of
    species to their environments. The effects of natural selection
    on genetic variation are complex, and can be difficult to disentangle
    from other evolutionary processes such as mutation, recombination,
    and genetic drift. The impact of selection on genetic variation
    is often studied using population genetic models, which can be
    used to generate predictions about the patterns of genetic variation

    % simulation
    As the complexity of population genetic datasets has increased,
    simulation has become a critical tool for interpretation, 
    analysis, and exploration of realistic evolutionary models.
    Simulation in population genetics has a long history 
    including both backward in time coalescent simulations
    \citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
    and forward in time simulations of complex demography and selection
    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.
    The development of simulation tools has been driven by the need to
    understand the effects of complex evolutionary processes on genetic
    variation, to provide a null model for hypothesis testing, to
    to explore the power and limitations of statistical methods, and 
    increasingly to provide a basis for machine learning and other
    simulation-based inference methods.

    % stdpopsim
    A lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers. This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community.
    While that is so, the current version of \stdpopsim does not include
    models of selection, which is a major limitation for many applications
    of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding the processes such
    as adaptation, the effects of selective sweeps, and the impact of
    background selection on genetic diversity. Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    accessible to a wide range of researchers and highly reproducible. 
    Further the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the density of functional elements, and relevant
    population size histories.

\section*{outline}
    \label{outline}
    \begin{enumerate}
        \item Single-population demographic inference methods
        \begin{enumerate}
            \item mscm2
            \item stairwaiplot
            \item GONE
            \item SMC++
        \end{enumerate}
        \item Multi-population demographic inference methods
        \begin{enumerate}
            \item dadi
            \item fastsimcoal
            \item momi2
        \end{enumerate}
        \item DFE inference methods
        \begin{enumerate}
            \item dadi
            \item polyDFE
            \item DFE-alpha
            \item GRAPES
        \end{enumerate} 
        \item sweeps
        \begin{enumerate}
            \item compare effect of BGS on sweep detection
            \item compare power in different pops
            \item compare methods
        \end{enumerate}
    \end{enumerate}
